---
title: "Using the R Package imager for Image Processing"
description: "A basic introduction to using the imager package."
date: "2026-01-01"
date-modified: "2026-01-01"
draft: false
lang: en
categories: [r]
---

::: {.callout-note}
Original Japanese version: [画像処理のためのRパッケージimagerの使い方](../../../posts/2025-01-01-imager/index.qmd)
:::

# Overview

When I wanted to do image analysis in R, I learned that there is a package called `{imager}`.
This post records what I tried.

# Loading the Package

Installation can be done normally with `install.packages()`.

```{r eval=FALSE}
install.packages("imager")
```

Load the package as follows.

```{r message=FALSE}
library("imager")
```

# Loading and Displaying Images

Use `load.image()` to load an image.
The package provides five sample images and one video.
You can check the samples as follows.

```{r}
list.files(system.file('extdata', package = 'imager'))
```

There is also a sample image of boats, but it does not seem to be stored here.

By the way, `system.file()` returns where a specified package is stored on the computer.
With `package = 'imager'`, the package is specified, and the first argument, `extdata`, specifies the location of the extra data.

Here, I load the sample parrot image and display it.
The usual `plot()` function is enough for display.

```{r}
file <- system.file('extdata/parrots.png', package = 'imager')
img <- load.image(file)
plot(img)
```

Use `load.dir()` to load multiple images at once.

```{r}
dir <- system.file('extdata/', package = 'imager')
list_img <- load.dir(dir)
plot(list_img)
```

When displaying your own image rather than a sample, specify the appropriate path and it should load correctly.
The same applies to `load.dir()`.

```r
file <- "/imagefile/hoge.jpg"
img <- load.image(file)
```

There are also dedicated functions for loading sample images.

```{r}
# sketch of birds by Leonardo, from Wikimedia
plot(load.example("birds"))

# Parotts from Kodak
plot(load.example("parrots"))

# The "coins" image comes from scikit-image.
plot(load.example("coins"))

# The Hubble Deep field (hubble) is from Wikimedia.
plot(load.example("hubble"))

# Boats from Kodak
plot(boats)
```

Only `boats` can be loaded directly.
When you want to try something quickly, `boats` is convenient.

# Checking Image Information

Images are stored in the `cimg` format.
If you type the object directly in the console, basic information is printed.

```{r}
boats
```

`boats` is a normal image with width 256 pixels, height 384 pixels, depth 1, and 3 channels.

Here, depth refers to frames.
For a video, images appear as a sequence, so the depth corresponds to the number of frames.

The number of channels is 3 for a color image, meaning RGB, and 1 for a grayscale image.

# Grayscale Conversion

Grayscale conversion can be done easily with `grayscale()`.

The second argument is `method`.
The default is `"Luma"`, which converts to grayscale using a linear approximation of luminance.
Another option is `"XYZ"`, in which case the image is assumed to be in the sRGB color space and CIE luminance is used.
If you do not have a particular reason to choose otherwise, leaving it omitted is fine.

The third argument is surprisingly important.
It is `drop = TRUE`.
When it is `TRUE`, the image is output with one channel.
When it is `FALSE`, the image is converted to grayscale while keeping three channels.
If you want to add color later, `FALSE` may be useful.
In most cases, the default should be fine.

```{r}
img_gray <- grayscale(boats)
layout(matrix(c(1, 2), 1, 2, byrow = TRUE))
plot(boats, main = "Original")
plot(img_gray, main = "Grayscale")
```
