---
title: "Drawing a Map of Japan in R"
description: "How to draw a map of Japan in R using the rnaturalearth and rnaturalearthhires packages."
date: "2026-01-31"
date-modified: "2026-01-31"
draft: false
lang: en
categories: [r]
---

::: {.callout-note}
Original Japanese version: [Rで日本地図を描く](../../../posts/2026-01-31-r-japan-map/index.qmd)
:::

## Installing and Loading Packages

```{.r}
install.packages("rnaturalearth")
```

If you use renv, install it with the following command.

```{.r}
renv::install("necountries")
```

Then load the package.

```{r}
library(rnaturalearth)
```

## Drawing a Map of Japan

To draw the map, the rnaturalearthdata package is also required.
The following code obtains and draws a map of Japan.

```{.r}
install.packages("rnaturalearthdata")
renv::install("rnaturalearthdata") # when using renv
```

It should work without explicitly loading the library, but I load it just in case.

```{r}
library(rnaturalearthdata)
```

Use the `ne_countries()` function to obtain map data for Japan, then draw it with `plot()`.

```{r}
japan <- ne_countries(
  scale = "medium",
  country = "Japan",
  returnclass = "sf"
)
plot(japan["geometry"])
```

The arguments are set as follows.

- `scale = "medium"`: obtains medium-resolution map data. `small` and `large` are also available.
- `country = "Japan"`: specifies map data for Japan.
- `returnclass = "sf"`: returns the map data as an sf object.

This lets you draw a map of Japan using R.
Customize the map style and details as needed.

::: {.callout-note}
## Official Documentation

For details on the `ne_countries()` function, see the official documentation.

- [Get natural earth world country polygons - ne_countries • rnaturalearth](https://ropensci.github.io/rnaturalearth/reference/ne_countries.html)
:::

## Drawing a Prefecture-Level Map of Japan

To draw a prefecture-level map of Japan, use the `rnaturalearthhires` package.
Install it from GitHub or R-universe.

::: {.callout-note}
## Why It Cannot Be Installed from CRAN

It seems that the package cannot be installed from CRAN because its size exceeds the recommended CRAN limit.
:::

```{.r}
remotes::install_github("ropensci/rnaturalearthhires")
```

Or install it with the following code.

```{.r}
install.packages(
  "rnaturalearthhires",
  repos = "https://ropensci.r-universe.dev",
  type = "source"
)
```

If you use renv, install it with the following command.

```{.r}
renv::install("ropensci/rnaturalearthhires")
```

Load the package.

```{r}
library(rnaturalearthhires)
```

Obtain a prefecture-level map of Japan and draw it.

```{r}
japan_prefectures <- ne_states(
  country = "Japan",
  returnclass = "sf"
)
plot(japan_prefectures["name_ja"])
```

The coastline is drawn neatly as well.

::: {.callout-note}
## Official Documentation

For details on the `ne_states()` function, see the official documentation.

- [Get natural earth world state (admin level 1) polygons - ne_states • rnaturalearth](https://ropensci.github.io/rnaturalearth/reference/ne_states.html)
:::

## References

- [High Resolution World Vector Map Data from Natural Earth used in rnaturalearth • rnaturalearthhires](https://docs.ropensci.org/rnaturalearthhires/index.html)
- [rnaturalearthhires - GitHub](https://github.com/ropensci/rnaturalearthhires)
