---
title: "Using the WFO Plant List API from R"
description: "Use the World Flora Online Plant List API from R to retrieve accepted names from scientific names."
date: "2026-06-16"
date-modified: "2026-06-16"
draft: false
lang: en
categories: [r, api, taxonomy]
execute:
  eval: false
---

::: {.callout-note}
Original Japanese version: [WFO Plant List APIをRから利用する方法](../../../posts/2026-06-16-wfo-r/index.qmd)
:::

[The World Flora Online](https://www.worldfloraonline.org/) is a global database of plant information.
It also provides an API as the [WFO Plant List API](https://list.worldfloraonline.org/), so this article shows how to access it from R and retrieve accepted names from scientific names.

## Installing Packages

Install the `httr2` package to make API requests easier.
Because API responses are returned as JSON, also install the `jsonlite` package.

After installation, load the packages with `library()`.

```{r}
# install.packages(c("httr2", "jsonlite"))
# renv::install(c("httr2", "jsonlite")) # renvを使用している場合はこちら
library(httr2)
library(jsonlite)
```

## Specifying the Endpoint

The API endpoint is the following URL.

```{r}
endpoint <- "https://list.worldfloraonline.org/gql.php"
```

## Creating a Query to Search by Scientific Name

The search is written as a GraphQL query.
Here, I create a query that searches for the scientific name `"Quercus serrata"` (called Konara in Japanese).

```{r}
query <- '
query {
  taxonNameSuggestion(
    termsString: "Quercus serrata"
    limit: 10
  ) {
    id
    fullNameStringHtml
    currentPreferredUsage {
      hasName {
        id
        fullNameStringHtml
      }
    }
  }
}
'
```

This query does the following.

- Calls the `taxonNameSuggestion` field
- Sets the scientific name to search for in the `termsString` argument
- Sets the maximum number of search results in the `limit` argument
- Retrieves `id`, `fullNameStringHtml`, and `currentPreferredUsage` fields as the query result

`fullNameStringHtml` is the searched scientific name expressed in HTML format.
`currentPreferredUsage` shows which taxon the name is currently treated as.
The `hasName` field inside `currentPreferredUsage` contains the scientific name of the current taxon.
In other words, when an accepted name exists, that name appears in the `hasName` field inside `currentPreferredUsage`.

## Creating and Sending the Request

Use `httr2::request()` to create the base request for the API.

```{r}
req <- request(endpoint)
```

Add the GraphQL `query` as JSON.

```{r}
req <- req_body_json(
  req,
  list(query = query),
  auto_unbox = TRUE
)
```

Then send it to WFO.

```{r}
resp <- req_perform(req)
```

Convert the returned JSON into an R `list`.
This makes the API response usable from R.

```{r}
x <- resp_body_json(resp, simplifyVector = FALSE)
```

::: {.callout}
## Using the Pipe Operator

The native pipe operator `|>` in base R can make the code easier to read.

```{r}
resp <- request(endpoint) |>
  req_body_json(
    list(query = query),
    auto_unbox = TRUE
  ) |>
  req_perform()
```

This pipe operator was introduced in R 4.1.0. It passes the object on the left as the first argument of the function on the right.
The `%>%` magrittr pipe operator, originally advocated by Hadley Wickham, was widely used before this, but with the introduction of R's native pipe operator, using `|>` is now often recommended.

The following documentation is useful for understanding the pipe operator.

- [Differences between the base R and magrittr pipes](https://tidyverse.org/blog/2023/04/base-vs-magrittr-pipe/)
:::

## Checking the Result

The `x` object stores the API response as an R list.

```{r}
print(x)
```

::: {.callout collapse="true"}

## Output Excerpt

```{.r}
$data
$data$taxonNameSuggestion
$data$taxonNameSuggestion[[1]]
$data$taxonNameSuggestion[[1]]$id
[1] "wfo-0000293164"

$data$taxonNameSuggestion[[1]]$fullNameStringHtml
[1] "<span class=\"wfo-name-full\" ><span class=\"wfo-name\"><i>Quercus</i> <i>serrata</i></span> <span class=\"wfo-name-authors\" >Murray</span></span>"

$data$taxonNameSuggestion[[1]]$currentPreferredUsage
$data$taxonNameSuggestion[[1]]$currentPreferredUsage$hasName
$data$taxonNameSuggestion[[1]]$currentPreferredUsage$hasName$id
[1] "wfo-0000293164"

$data$taxonNameSuggestion[[1]]$currentPreferredUsage$hasName$fullNameStringHtml
[1] "<span class=\"wfo-name-full\" ><span class=\"wfo-name\"><i>Quercus</i> <i>serrata</i></span> <span class=\"wfo-name-authors\" >Murray</span></span>"




$data$taxonNameSuggestion[[2]]
$data$taxonNameSuggestion[[2]]$id
[1] "wfo-0000293165"
...
```
:::

Ten candidates were returned because the query specified `limit: 10`.
Let's inspect the first result.

```{r}
x$data$taxonNameSuggestion[[1]]
```

::: {.callout collapse="true"}

## Output

```{.r}
$id
[1] "wfo-0000293164"

$fullNameStringHtml
[1] "<span class=\"wfo-name-full\" ><span class=\"wfo-name\"><i>Quercus</i> <i>serrata</i></span> <span class=\"wfo-name-authors\" >Murray</span></span>"

$currentPreferredUsage
$currentPreferredUsage$hasName
$currentPreferredUsage$hasName$id
[1] "wfo-0000293164"

$currentPreferredUsage$hasName$fullNameStringHtml
[1] "<span class=\"wfo-name-full\" ><span class=\"wfo-name\"><i>Quercus</i> <i>serrata</i></span> <span class=\"wfo-name-authors\" >Murray</span></span>"
```

:::

The three fields `id`, `fullNameStringHtml`, and `currentPreferredUsage` were returned.

The field to focus on is `hasName` inside `currentPreferredUsage`.
This field contains the scientific name of the current taxon.

```{r}
x$data$taxonNameSuggestion[[1]]$currentPreferredUsage
```

::: {.callout}

## Output

```{.r}
$hasName
$hasName$id
[1] "wfo-0000293164"

$hasName$fullNameStringHtml
[1] "<span class=\"wfo-name-full\" ><span class=\"wfo-name\"><i>Quercus</i> <i>serrata</i></span> <span class=\"wfo-name-authors\" >Murray</span></span>"
```

:::

In this example, the `hasName` field inside `currentPreferredUsage` displays the scientific name "*Quercus serrata* Murray".

```{r}
x$data$taxonNameSuggestion[[1]]$id
x$data$taxonNameSuggestion[[1]]$currentPreferredUsage$hasName$id
```

::: {.callout collapse="true"}

## Output

```{.r}
[1] "wfo-0000293164"
[1] "wfo-0000293164"
```

:::

Comparing `x$data$taxonNameSuggestion[[1]]$id` with `x$data$taxonNameSuggestion[[1]]$currentPreferredUsage$hasName$id`, the two values are identical.

This shows that the searched scientific name `"Quercus serrata"` is currently treated as `"Quercus serrata"`.

In this way, you can retrieve an accepted name from a searched scientific name.

## Creating a Function to Retrieve Accepted Names

In practical work, it is useful to wrap the process in functions.

Here, I turn the steps above into functions and create a function that retrieves the accepted name from a scientific name.
First, create `get_wfo_suggestions()`, which retrieves candidates from a scientific name.

In addition to the query above, this function also retrieves fields such as `fullNameStringNoAuthorsPlain`, `authorsString`, and `rank`.
The `rank` field is especially important because it shows the taxonomic rank of the searched name.
For example, it lets you distinguish whether the name is at the rank of `species`, `subspecies`, `variety`, and so on.

When retrieving an accepted name, it is often useful to look for an accepted name at the same rank as the searched name, so this function also retrieves rank information.

```{r}
get_wfo_suggestions <- function(
  name,
  limit = 10,
  endpoint = "https://list.worldfloraonline.org/gql.php"
) {
  query <- '
  query NameSearch($terms: String!, $limit: Int) {
    taxonNameSuggestion(
      termsString: $terms
      limit: $limit
    ) {
      id
      fullNameStringPlain
      fullNameStringNoAuthorsPlain
      authorsString
      rank
      currentPreferredUsage {
        hasName {
          id
          fullNameStringPlain
          fullNameStringNoAuthorsPlain
          authorsString
          rank
        }
      }
    }
  }
  '

  req <- request(endpoint)

  req <- req_body_json(
    req,
    list(
      query = query,
      variables = list(
        terms = name,
        limit = limit
      )
    ),
    auto_unbox = TRUE
  )

  resp <- req_perform(req)
  x <- resp_body_json(resp, simplifyVector = FALSE)

  res <- x$data$taxonNameSuggestion

  if (is.null(res) || length(res) == 0) {
    return(data.frame())
  }

  get_value <- function(z, field) {
    if (is.null(z[[field]])) {
      NA_character_
    } else {
      z[[field]]
    }
  }

  get_accepted <- function(z, field) {
    if (is.null(z$currentPreferredUsage)) {
      NA_character_
    } else if (is.null(z$currentPreferredUsage$hasName[[field]])) {
      NA_character_
    } else {
      z$currentPreferredUsage$hasName[[field]]
    }
  }

  out <- data.frame(
    input = name,
    id = sapply(res, get_value, field = "id"),
    name = sapply(res, get_value, field = "fullNameStringPlain"),
    name_no_author = sapply(
      res,
      get_value,
      field = "fullNameStringNoAuthorsPlain"
    ),
    authors = sapply(res, get_value, field = "authorsString"),
    rank = sapply(res, get_value, field = "rank"),
    accepted_id = sapply(res, get_accepted, field = "id"),
    accepted_name = sapply(res, get_accepted, field = "fullNameStringPlain"),
    accepted_name_no_author = sapply(
      res,
      get_accepted,
      field = "fullNameStringNoAuthorsPlain"
    ),
    accepted_authors = sapply(res, get_accepted, field = "authorsString"),
    accepted_rank = sapply(res, get_accepted, field = "rank"),
    stringsAsFactors = FALSE
  )

  out$is_accepted <- !is.na(out$accepted_id) & out$id == out$accepted_id

  out
}
```

Next, create `get_accepted_name()`, a function that retrieves the accepted name from a scientific name.

```{r}
get_accepted_name <- function(name, rank = "species") {
  tab <- get_wfo_suggestions(name)

  if (nrow(tab) == 0) {
    return(NA)
  }

  hit <- tab[
    tab$rank == rank &
      tab$name_no_author == name,
  ]

  if (nrow(hit) == 0) {
    return(NA)
  }

  hit[1, ]
}
```

The `get_accepted_name()` function takes a scientific name and rank, then calls `get_wfo_suggestions()` to retrieve candidates.

```{r}
result <- get_wfo_suggestions("Quercus serrata")
print(result)
```

::: {.callout collapse="true"}

## Output

```{.r}
             input             id
1  Quercus serrata wfo-0000293164
2  Quercus serrata wfo-0000293165
3  Quercus serrata wfo-0001062616
4  Quercus serrata wfo-0000293167
5  Quercus serrata wfo-0000293168
6  Quercus serrata wfo-0000293169
7  Quercus serrata wfo-0000809955
8  Quercus serrata wfo-0000804579
9  Quercus serrata wfo-0000804580
10 Quercus serrata wfo-1200020125
                                                      name
1                                   Quercus serrata Murray
2                                    Quercus serrata Roxb.
3                                   Quercus serrata Thunb.
4                     Quercus serrata var. attenuata Blume
5        Quercus serrata var. brevipetiolata (A.DC.) Nakai
6                     Quercus serrata var. chinensis Wenz.
7              Quercus serrata f. concolor (Sugim.) H.Ohba
8                     Quercus serrata var. concolor Sugim.
9  Quercus serrata var. donarium (Nakai) Kitam. & T.Horik.
10                   Quercus serrata var. longicarpa Uyeki
                        name_no_author                   authors    rank
1                      Quercus serrata                    Murray species
2                      Quercus serrata                     Roxb. species
3                      Quercus serrata                    Thunb. species
4       Quercus serrata var. attenuata                     Blume variety
5  Quercus serrata var. brevipetiolata             (A.DC.) Nakai variety
6       Quercus serrata var. chinensis                     Wenz. variety
7          Quercus serrata f. concolor           (Sugim.) H.Ohba    form
8        Quercus serrata var. concolor                    Sugim. variety
9        Quercus serrata var. donarium (Nakai) Kitam. & T.Horik. variety
10     Quercus serrata var. longicarpa                     Uyeki variety
      accepted_id                              accepted_name
1  wfo-0000293164                     Quercus serrata Murray
2  wfo-0000814085 Castanopsis indica (Roxb. ex Lindl.) A.DC.
3            <NA>                                       <NA>
4  wfo-0000289386                Quercus acutissima Carruth.
5  wfo-0000807317            Quercus serrata subsp. serrata
6  wfo-0000293817                   Quercus variabilis Blume
7  wfo-0000293164                     Quercus serrata Murray
8  wfo-0000293164                     Quercus serrata Murray
9  wfo-0000807317            Quercus serrata subsp. serrata
10           <NA>                                       <NA>
          accepted_name_no_author        accepted_authors accepted_rank
1                 Quercus serrata                  Murray       species
2              Castanopsis indica (Roxb. ex Lindl.) A.DC.       species
3                            <NA>                    <NA>          <NA>
4              Quercus acutissima                Carruth.       species
5  Quercus serrata subsp. serrata                    <NA>    subspecies
6              Quercus variabilis                   Blume       species
7                 Quercus serrata                  Murray       species
8                 Quercus serrata                  Murray       species
9  Quercus serrata subsp. serrata                    <NA>    subspecies
10                           <NA>                    <NA>          <NA>
   is_accepted
1         TRUE
2        FALSE
3        FALSE
4        FALSE
5        FALSE
6        FALSE
7        FALSE
8        FALSE
9        FALSE
10       FALSE
```

:::

Use `get_accepted_name()` to retrieve the row that contains the accepted name for a scientific name.

```{r}
accepted_name <- get_accepted_name("Quercus serrata")
print(accepted_name)
```

::: {.callout collapse="true"}

## Output

```{.r}
            input             id                   name  name_no_author
1 Quercus serrata wfo-0000293164 Quercus serrata Murray Quercus serrata
  authors    rank    accepted_id          accepted_name
1  Murray species wfo-0000293164 Quercus serrata Murray
  accepted_name_no_author accepted_authors accepted_rank is_accepted
1         Quercus serrata           Murray       species        TRUE
```

:::

## Notes on Using the API

The WFO Plant List API is publicly available, but sending a large number of requests in a short time may **place load on the server**.
Searching a small number of names for learning or checking is fine, but when processing many scientific names at once, keep the following points in mind.

* Save or cache results so that the same name is not queried repeatedly.
* Add `Sys.sleep()` in loops when needed to avoid continuous access.
* For large datasets, consider using the published download data or a local workflow instead of trying to retrieve all data through the API.
* Do not accept candidates returned by the API without review; check `rank`, WFO ID, `currentPreferredUsage`, and accepted name.
* Because `taxonNameSuggestion()` is for candidate search, make the candidate-review rules explicit when doing strict bulk matching.

For example, when running a loop, you can add `Sys.sleep()` as follows to avoid continuous access.

```{r}
for (nm in names) {
  result <- get_accepted_name(nm)
  Sys.sleep(0.5) # サーバー負荷を下げるため少し待つ
}
```

The WFO Plant List also provides the same data as the API through Zenodo, with a DOI that can be cited.
For research use cases that process many names or require reproducibility, consider not relying only on the API. Instead, specify the WFO Plant List release used and use or cite the data available on Zenodo.

Zenodo Link: <a href="https://doi.org/10.5281/zenodo.18007552"><img src="https://zenodo.org/badge/DOI/10.5281/zenodo.18007552.svg" alt="DOI"></a>
