---
title: "Creating Vectors in R"
description: "A summary of several ways to create vectors in R."
date: "2024-06-18"
date-modified: "2024-06-18"
draft: false
lang: en
categories: [r]
---

::: {.callout-note}
Original Japanese version: [Rでベクトルを作成する](../../../posts/2024-06-18-r-vector/index.qmd)
:::

This post summarizes several ways to create vectors in R.
The functions introduced here are:

- `c()`
- `:`
- `seq()`
- `seq_len()`
- `seq_along()`
- `rep()`

## `c()` and `:`

The simplest ways to create vectors in R are `c()` and `:`.

If you list values separated by commas inside `c()`, the result is a vector.
The name `c` comes from "combine."

```{r}
c(1, 2, 3)
```

You can also create a vector by specifying a start and end point in the form `[start]:[end]`.

```{r}
1:3
```

These are used very frequently, so they are worth remembering.

## `seq()`

The `seq()` function comes from "sequence." As the name suggests, it arranges values according to the arguments supplied inside `()`.

The most basic use is to create a vector by specifying the first and last values.

```{r}
seq(1, 3) # seq(start, end)
```

```{r}
seq(from = 1, to = 3) # explicit form
```

If you specify only one value, `seq()` creates a vector that increases by 1 from 1 to that value.

```{r}
seq(3)
```

When the sequence starts from 1, `seq_len()` is faster.

```{r}
seq_len(5) # vector from 1 to 5
```

### Creating a Vector with a Specified Step Size

By default, the value increases by 1, but you can change the step size with the `by` argument.

```{r}
seq(from = 1, to = 3, by = 0.5) # increase by 0.5
```

### Creating a Vector with a Specified Length

If you specify the `length.out` argument, you can create a vector with a specified length.
R divides the range between the first and last values into equal intervals.

```{r}
seq(from = 1, to = 3, length.out = 6) # vector of length 6
```

Use `seq_along()` to create a vector with the same length as another vector.

```{r}
x <- seq_along(c("A", "B", "C", "D", "E"))
seq_along(x)
```

This is useful when writing `for` loops.
It is common to write a `for` loop like this:

```{r}
x <- c("A", "B", "C", "D", "E")
for (i in 1:length(x)) {
  print(x[i])
}
```

Using `seq_along()`, it can be rewritten as follows.

```{r}
x <- c("A", "B", "C", "D", "E")
for (i in seq_along(x)) {
  print(x[i])
}
```

There is not much visible difference, but the behavior changes when the vector used in the loop has length 0, that is, when it is an empty vector.

## `rep()`

Use the `rep()` function to create a vector by repeating the same value.
The name comes from "repeat."

```{r}
rep(3, 5) # repeat 3 five times
```

You can also repeat a vector.

```{r}
rep(1:3, 3) # repeat 1 to 3 three times
rep(1:3, times = 3) # same as above
```

Each value can also be repeated as a block.

```{r}
rep(1:3, each = 3) # repeat each value three times
```

You can specify the number of repeats for each value.

```{r}
rep(1:3, c(1, 2, 1)) # 1 once, 2 twice, 3 once
```

You can also specify the length.
Because the length takes priority, the repetition is truncated if needed.

```{r}
rep(1:3, length.out = 5) # repeat up to length 5
```

## Examples

These vector creation methods are useful when creating IDs or groups in data frames, or when interpolating points for plots.
For example, if you want to assign groups A to C to rows in a data sheet, `rep(c("A", "B", "C"), each = 5)` makes that easy.

If you want to draw a function such as a sine curve, you can do it simply as follows.

```{r}
x <- seq(from = 0, to = pi, length.out = 100)
plot(x, sin(x))
```

## Summary

Personally, I use `c()` and `:` almost every day, while I use the `seq()` family and `rep()` only occasionally.
Because of that, when I suddenly need a long vector, I sometimes cannot produce it immediately.
It is easy to look up, but it is useful to be able to create the needed vector quickly.
