---
title: "Removing Duplicate Vector Elements in R"
description: "A note on using the `make.unique()` function to make duplicate vector elements unique."
date: "2026-01-27"
date-modified: "2026-01-27"
draft: false
lang: en
categories: [r]
---

::: {.callout-note}
Original Japanese version: [Rでベクトルの重複を取り除く方法](../../../posts/2026-01-27-make-unique/index.qmd)
:::

## Using the `make.unique()` Function

This is a note on how to handle duplicate elements in a vector in R and turn them into unique elements.
With the `make.unique()` function, duplicate elements can be automatically converted into unique values by adding suffixes.
The usage is as follows.

```{.r}
make.unique(names, sep = ".")
```

- `names`: a character vector you want to make unique
- `sep`: the separator used before the suffix added to duplicate elements; the default is a dot, `.`

## Example

Suppose we have the following vector.

```{r}
x <- c("apple", "banana", "apple", "orange", "banana")
```

Applying `make.unique()` to this vector gives the following result.

```{r}
unique_x <- make.unique(x)
print(unique_x)
```

Because `apple` and `banana` are duplicated, `.1` is added after the repeated words.

## Closing Note

I cannot remember exactly where I used this function, but I remembered using it before and left this note.
The opportunities to use it may be limited, but it can be useful to know.
